---
title: "Kaggle Poker Rule Induction"
author: "June Lau"
date: "Monday, December 29, 2014"
output: pdf_document
---

# Purpose
To document a KNN model approach to inferring rules for the Kaggle challenge <a href="https://www.kaggle.com/c/poker-rule-induction"> Poker Rule Induction </a>. Note that the given train dataset was not splitinto a training and test set, as is standard practice for machine learning. This document is part 2.

To begin, we navigate to the required directory and load up the required libraries.


## Required libraries
```{r}
setwd("O:/random projects/poker-rule-induction/")
#libraries for naive bayes
library("plyr")
library('klaR')
library('caret')
```

# Data pre-processing
Next we import the datasets and have a look at a summary and a few random rows.
```{r}
train <- read.table(unz("train.csv.zip", "train.csv"), header=T, sep =",")
test <- read.table(unz("test.csv.zip", "test.csv"), header=T, sep =",")
train <-data.frame(train)
# summaries
count(train,"hand")
str(train)
```


##  Method 2: K nearest neighbour (kNN)
Given the Naive Bayes' poor performance at guessing the poker hand given a set of 5 cards, let's look at another common classifier - the k-nearest neighbour method. Hopefully this will take into account the interaction between the input variables - e.g. cards of the same suit, or cards in a series.

We're using the same form of train as we did for Naive Bayes. That is, all columns are recognised as factors. 

```{r}
# load up required library
library ("class")
# class labels is 'hand' variable
cls <- train[,11]

# apply the model to the test dataset. We exclude column 11 from the test dataset, as this is what we wish to predict
model.knn <- knn(train[,-11], test[,-11], cls, k = 1)
summary(model.knn)
```


Since I haven't worked out how to evaluate a kNN model, we'll look at some sample rows from the test dataset.
```{r}

# how do you evaluate a knn model?
cbind(test[10,],model.knn[10]) # incorrect - should be 0
cbind(test[152,],model.knn[152]) # two pairs. Incorrect as hand=8 is a straight flush
```

Again, this doesn't look good. Uploading to kaggle confirms this :(